---
title: "HW2"
author: "Xiaofu Dai"
date: "February 16, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part 1
```{r}
library(ggplot2)
library(viridis)
palette(viridis(256))
library(future.apply)
plan(multicore)
```
### function
I use f(x) = x^5 - x^2 +1, f(x) = x^6 + 1/x - 5 to plot the art show plots.
```{r}
# Functions
F1 <- function(x){ 
  fx = x^5 - x^2 +1
  dfx = 5*x^4-2*x 
  return(c(fx, dfx)) 
}

F2 <- function(x){ 
  fx = x^6 + 1/x - 5
  dfx = 6*(x^5) - 1/(x^2)
  return(c(fx, dfx)) 
}

# global control variable 
bRootOrIterations<-0 

# Here's the function that performs Newton-Raphson
TwoDNewtonRaphson<-function(func,StartingValue,Tolerance,MaxNumberOfIterations){
  i<-0  
  NewZ<- StartingValue  
  Deviation=abs(func(StartingValue)[1])   

  while ((i<MaxNumberOfIterations)&&(Deviation>Tolerance))
  {
    Z<-func(NewZ)   
    if ((Z[1]=="NaN")||(Z[2]=="NaN")){
      cat("Function or derivative not defined error.\n"); break
    }
    NewZ <- NewZ-Z[1]/Z[2]
    NewVal <- func(NewZ)
    Deviation <- abs(NewVal[1])
    i<-i+1
  }

  # output the result
  if (Deviation>Tolerance){
    cat(paste("\nConvergence failure. Deviation:",Deviation, "after ", i, 	"iterations"))
  }

  if (bRootOrIterations==1){
   return(NewZ)
  }else{
    return(c(i,i))
  }
}



# A function to check whether two points are close together
CloseTo<-function(x,y){
  # returns 1 if x is close to y
  if (abs(x-y)<0.1) {
    return(1)
  }else{
    return (0)
  }
}

# do plot
Root_calculator <- function(Funcn, xmin, xmax, xsteps, ymin, ymax, ysteps) {
  
  x <- seq(xmin, xmax, length.out = xsteps)
  y <- seq(ymin, ymax, length.out = ysteps)

  out_dat <- expand.grid(x = x, y = y)

  ThisZ <- complex(1, out_dat$x, out_dat$y)

  Root <- sapply(ThisZ,
                 FUN = TwoDNewtonRaphson,
                 func = Funcn,
                 Tolerance = 1e-1,
                 MaxNumberOfIterations = 100)

  if(bRootOrIterations == 0) {
    out_dat$color <- 104 + 5 * Root[1, ]
    out_dat$root1 <- Root[1, ]
    out_dat$root2 <- Root[2, ]
  } else {
    out_dat$color <- 5 * Root
    out_dat$root1 <- Root
  }

  return(out_dat)
}



Func1 <- Root_calculator(F1, -1/4, 1/4, 200, -1/4, 1/4, 200)
ggplot(Func1, aes(x, y,color = root2)) +
  geom_point()+
    theme_minimal() +
   scale_color_gradient(low="pink", high = "white")+
    coord_fixed() +
    labs(title = "A Midspring Night's Dream")



Func2 <- Root_calculator(F2, -1/2, 1/2, 200, -1/2, 1/2, 200)
ggplot(Func2, aes(x, y, fill = root2)) +
    geom_tile()+
    scale_color_gradient2(high="blue", mid="pink",
                     low="blue", space ="Lab" ) +
    theme_minimal() +
    coord_fixed() +
    labs(title = "Crystal Mandala")

```

## Part 2
1.Write a program to implement the Secant method. 
2.Test it, using x0=1 and x1=2 on: 
   1.cos(x)-x 
   2.log(x)-exp(-x) 
Compare it with the performance of Newton-Raphson on the same functions. 
3. Write a function to show a plot of the iterations of the algorithm.


Compare Newton's Method with the secant method, we see Newton's method requires the evaluation of both f(x) and its derivative f'(x) at every step, while the secant method only requires the evaluation of f(x). So if it's much easier to calculate  f(x) or we can't get f'(x), the secant method is a better choice.

1. Secant method
```{r}
# Functions
f1 = function(x){
  return(cos(x)-x)
}
f2 = function(x){
  return(log(x)-exp(-x))
}

# a sufficiently small difference between xn and xn-1
secant = function(func, x0, x1, Tolerance, MaxNumberOfIterations){
  i=0
  while (abs(x1-x0) > Tolerance & i < MaxNumberOfIterations) {
    x2 = x1 - func(x1)*(x1-x0)/(func(x1)-func(x0))
    x0 = x1
    x1 = x2
    i = i+1
  }
  return(c(i,x1))
}
```

2. Newton-Raphson
```{r}
# Functions
f3 <- function(x){ 
  fx = cos(x)-x
  dfx = -sin(x)-1
  return(c(fx, dfx)) 
}

f4 <- function(x){ 
  fx = log(x)-exp(-x)
  dfx = 1/x + exp(-x)
  return(c(fx, dfx)) 
}

# global control variable 
bRootOrIterations<-0 

# Here's the function that performs Newton-Raphson
NewtonRaphson<-function(func,StartingValue,Tolerance,MaxNumberOfIterations){
  i<-0  
  NewZ<- StartingValue  
  Deviation=abs(func(StartingValue)[1])   

  while ((i<MaxNumberOfIterations)&&(Deviation>Tolerance))
  {
    Z<-func(NewZ)   
    if ((Z[1]=="NaN")||(Z[2]=="NaN")){
      cat("Function or derivative not defined error.\n"); break
    }
    NewZ <- NewZ-Z[1]/Z[2]
    NewVal <- func(NewZ)
    Deviation <- abs(NewVal[1])
    i<-i+1
  }

  # output the result
  if (Deviation>Tolerance){
    cat(paste("\nConvergence failure. Deviation:",Deviation, "after ", i, 	"iterations"))
  }

  if (bRootOrIterations==1){
   return(NewZ)
  }else{
    return(c(i,NewZ))
  }
}
```
Compare it with the performance of Newton-Raphson on the same functions. 
```{r}
# parameters 
x0 = 1
x1 =2 

# for 1. cos(x)-x
# Secant Method
secant(f1, x0, x1, 1e-04, 40) 
# Newton Method
NewtonRaphson(f3, x0, 1e-04, 40)

# for 2.log(x)-exp(-x)
# Secant Method
secant(f2, x0, x1, 1e-04, 40) 
# Newton Method
NewtonRaphson(f4, x0, 1e-04, 40)


# Two methods got similar roots for two functions, separately.
# But the Newton's method took less interations to get the root than that of the secant method.
# The Newton's method took 2 and 3 interations to get the root of function 1 and function 2, respectively, while secant method took 4 and 5 interactions, respectively. 

# Compare Newton's method with the secant method, we see that Newton's method converges faster. And as mentioned before, Newton's method requires the evaluation of both f(x) and its derivative f'(x) at every step, while the secant method only requires the evaluation of f(x). 
```
Write a function to show a plot of the iterations of the algorithm.
```{r}
# For Secant Method
library(shape)
secantplot = function(func, x0, x1, Tolerance, MaxNumberOfIterations,DrawLines)
  {
  i=0
  x0 = x0
  x1 = x1
  
  while (abs(x1-x0) > Tolerance & i < MaxNumberOfIterations) 
    {
    y0 = func(x0)
    y1 = func(x1)
    x2 = x1 - func(x1)*(x1-x0)/(func(x1)-func(x0))  # find another point on xaxis
    
    if (DrawLines){
			Arrows(x0,0,x0,y0,col="green",lty=2,arr.length=0.25)
      Arrows(x1,0,x1,y1,col="green",lty=2,arr.length=0.25)
			Arrows(x0,y0,x2,0,col="blue",lty=2, arr.length=0.25)
			}
    
    x0 = x1
    x1 = x2
    i = i+1
  }
}

# 1. cox(x)-x
curve(cos(x)-x, main="Secant's: cos(x)-x", xlim = c(0.5,2), ylim = c(-2.5,0))
abline(h = 0, v = 0)
secantplot(f1, x0, x1, 1e-04, 40,1)

# 2.log(x)-exp(-x)
curve(log(x)-exp(-x), main="Secant's: log(x)-exp(-x)", xlim = c(1,2), ylim = c(-0.5,0.6))
abline(h = 0, v = 0)
secantplot(f2, x0, x1, 1e-04, 40, 1)
```

```{r}
# For Newton Method
FixedPointFinder<-function(func,StartingValue,Tolerance,MaxNumberOfIterations,DrawLines){
	
	i<-0
	Xprime<-StartingValue 
	Dev = 1000
	
	while ((i<MaxNumberOfIterations)&&(Dev>Tolerance))
	{
		
		X<-Xprime     # x axis
		Z<-func(X)    # f(x) and f'(x)
		Y<-Z[1]       # y
		k<-Z[2]       # f'(x)
		if ((Z[1]=="NaN")||(Z[2]=="NaN")){
      cat("Function or derivative not defined error.\n"); break
		}
		
		X2 = X-Y/k
		
		if (DrawLines){
			Arrows(X,0,X,Y,col="blue",lty=2,arr.length=0.25)
			Arrows(X,Y,X2,0,col="blue",lty=2,arr.length=0.25)
			}
	
		Xprime = X2
		NewVal <- func(X2)
	  Dev <- abs(NewVal[1])
		i<-i+1
	}
	# output the result
	if (Dev<Tolerance){
		cat(paste("\nFound the fixed point: ",X, "after ", i, "iterations"))
		}else{
		cat(paste("\nConvergence failure. Deviation: ",Dev, "after ", i, 	"iterations"))}
}

# 1. cox(x)-x
curve(cos(x)-x, main="Newton's: cos(x)-x", ylim=c(-0.5, 0), xlim = c(0.7,1))
abline(h = 0, v = 0)
FixedPointFinder(f3, 1, 1e-04, 40,1)

# 2.log(x)-exp(-x)
curve(log(x)-exp(-x), main="Newton's: log(x)-exp(-x)", ylim=c(-0.4,0.1), xlim = c(1,1.4))
abline(h = 0, v = 0)
FixedPointFinder(f4, 1, 1e-04, 40,1)
```


